---
title: "Scrapped COde"
author: "DSW"
date: "2025-04-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


This is code we do not need but want to save just in case




# Initial Arima Model

```{r Initial Arima Model}

# simple model
daily_load_data_ts_simple <- ts(daily_load_data, frequency = 365)
#arima
model_auto_simple <- auto.arima(daily_load_data_ts_simple)
summary(model_auto_simple)


# decompose the series 
decompose_daily_load_data <- decompose(daily_load_data, type = "additive")
# plot the decomposed series
plot(decompose_daily_load_data)
# deseasonalized series
deseasonal_daily_load_data <- seasadj(decompose_daily_load_data)

# plot the deseasonalized series
autoplot(deseasonal_daily_load_data, 
         main = "Deseasonalized Daily Load Data", 
         ylab = "Load Value", xlab = "Time")

# auto.arima() 
model_auto_deseasonal <- auto.arima(deseasonal_daily_load_data)
summary(model_auto_deseasonal)
```

# Original Forecast

```{r Original Forecast}
# moving to 2010 data
daily_load_data_forecasting <- load_data %>% 
  filter(date < as.Date("2010-12-31")) %>%
  select(h1:h24) %>%
  rowSums(na.rm = TRUE) %>%
  msts(seasonal.periods =c(7,365.25), start = c(2005, 1, 1))

autoplot(daily_load_data_forecasting)

# model
decompose_daily_load_data_forecasting <- decompose(daily_load_data_forecasting, type = "additive")
plot(decompose_daily_load_data_forecasting)
deseasonal_daily_load_data_forecasting <- seasadj(decompose_daily_load_data_forecasting)
model_auto_deseasonal_forecasting <- auto.arima(deseasonal_daily_load_data_forecasting)
summary(model_auto_deseasonal_forecasting)


# forecast period (from January 1, 2011 to February 28, 2011)
forecast_start_date <- as.Date("2011-01-01")
forecast_end_date <- as.Date("2011-02-28")

# number of periods for the forecast
num_periods <- as.integer(forecast_end_date - forecast_start_date) + 1

forecast_deseasonal <- forecast(model_auto_deseasonal_forecasting, h = num_periods)

plot(forecast_deseasonal, 
     main = "Forecast for Deseasonalized Data (Jan 1, 2011 to Feb 28, 2011)", 
     ylab = "Deseasonalized Values", xlab = "Time")


```





```{r}
# Fixed Original Forecast

```{r Fixed OG Forecast}
n_for <- 12  # Forecast horizon (12 months)

# Create the `msts` time series object for daily load data with both weekly and yearly seasonalities
# Here we're assuming `daily_load_data` is your vector or numeric data of daily load values
daily_load_data_msts <- msts(
  daily_load_data[1:(length(daily_load_data) - n_for)],  # Exclude the last 12 months for training
  seasonal.periods = c(7, 365.25)  # Weekly (7 days) and Yearly (365.25 days) seasonalities
)

# Check the first 15 rows of the msts time series
head(daily_load_data_msts, 15)

# Check the last 15 rows of the msts time series
tail(daily_load_data_msts, 15)

# Decompose the 'daily_load_data_msts' using the 'additive' method
decompose_daily_load_msts <- decompose(daily_load_data_msts, type = "additive")

# Plot the decomposed series
plot(decompose_daily_load_msts)

MEAN_seas <- meanf(y = daily_load_data_msts, h = 12)

# Checking residuals of the forecast model
checkresiduals(MEAN_seas)

SNAIVE_seas <- snaive(daily_load_data_msts, h = 12)

# Checking residuals of the Seasonal Naive model
checkresiduals(SNAIVE_seas)

SARIMA_autofit <- auto.arima(daily_load_data_msts)

# Check the residuals of the SARIMA model
checkresiduals(SARIMA_autofit)

```
```

```

